# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data/burden-of-disease-cmnn.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.

# we will be working with Tanzania, Kenya, Singapore

data_cmnn_mini <-
  data_cmnn %>% 
  filter(Entity%in% c("Tanzania", 
                     "Kenya", 
                     "Singapore")) %>% 
  rename(DALYS = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`,
         Country = Entity)
  

```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
# Use kable() from the knitr package to render the table.

data_cmnn_wide <- 
  data_cmnn_mini %>%
  select(Country, Year, DALYS ) %>% 
  pivot_wider(names_from = Country,
              values_from = DALYS) %>% 
  kable()

data_cmnn_wide

```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.

cmnn_trend <- 
  data_cmnn_mini %>% 
  ggplot(mapping = aes(x = Year,
                       y = DALYS,
                       color = Country)) +
  geom_line() +
  labs (
    x = "Years",
    y = "DALYs lost",
    title = "Plot Showing Trends in CMNN Burden Over Time"
  )
  

cmnn_trend

```

## Summary of CMNN Burden Findings

Tanzania and Kenya had higher DALYs in early 1990s to 2000s due to HIV epidemic unlike Singapore which seems to have stable DALYs over the past 30 years.  

## Attribution

This section was completed by Ludovick L. Lopa

The link to the group's GitHub repository is https://github.com/Ludovick-Lopa/group2_dlp

