# Overall Disease Burden

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here, dplyr, ggplot2)

# Import the overall disease data
data_overall <- read_csv(here("data/burden-of-disease-all-causes.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_overall_mini <- data_overall %>% filter(Entity %in% c("Tanzania", "Kenya", "Singapore")) %>% 
  rename(DALYS = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`)
```

## Table of Estimates for Overall Disease Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.


data_overall_wide <- data_overall_mini%>%
  
  select(Entity, Year, DALYS ) %>% 
  pivot_wider(names_from = Entity,
              values_from = DALYS) %>% 
  kable()

data_overall_wide
 



# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.


ggplot(data_overall_mini, aes(x = Year, y = DALYS, color = Entity)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Trend of DALY Burden Over Time",
    x = "Year",
    y = "DALYs (Age-standardized rate)",
    color = "Country"
  ) +
  theme_minimal()



```

## Summary of Overall Disease Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.

## Attribution

This section was completed by [Daniel Kadoke!]

The link to the group's GitHub repository is [https://github.com/Ludovick-Lopa/group2_dlp!]
